taxize - taxonomic search and retrieval in R

### 2013-07-19 inst/doc/taxize_withcode_plos.tex
more work on comments

\documentclass[10pt]{article}\usepackage{graphicx, color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.2, 0.2, 0.2}
\newcommand{\hlnumber}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlfunctioncall}[1]{\textcolor[rgb]{0.501960784313725,0,0.329411764705882}{\textbf{#1}}}%
\newcommand{\hlstring}[1]{\textcolor[rgb]{0.6,0.6,1}{#1}}%
\newcommand{\hlkeyword}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
\newcommand{\hlargument}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hlcomment}[1]{\textcolor[rgb]{0.180392156862745,0.6,0.341176470588235}{#1}}%
\newcommand{\hlroxygencomment}[1]{\textcolor[rgb]{0.43921568627451,0.47843137254902,0.701960784313725}{#1}}%
\newcommand{\hlformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hleqformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hlassignement}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
\newcommand{\hlpackage}[1]{\textcolor[rgb]{0.588235294117647,0.709803921568627,0.145098039215686}{#1}}%
\newcommand{\hlslot}[1]{\textit{#1}}%
\newcommand{\hlsymbol}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlprompt}[1]{\textcolor[rgb]{0.2,0.2,0.2}{#1}}%

@@

First, we must install taxize. There are two versions of taxize, a) a stable release that can be installed from the R package repository, CRAN, and b) from GitHub \cite{github_taxize}, where the code is developed.

Installing from CRAN or GitHub

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{install.packages}(\hlstring{"taxize"})
\end{alltt}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{install.packages}(\hlstring{"devtools"})
\hlfunctioncall{require}(devtools)
\hlfunctioncall{install_github}(\hlstring{"taxize_"}, \hlstring{"ropensci"})
\end{alltt}
\end{kframe}
\end{knitrout}

Load taxize into your R session.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{library}(taxize)
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
temp <- \hlfunctioncall{gnr_resolve}(names = \hlfunctioncall{c}(\hlstring{"Helianthos annus"}, \hlstring{"Homo saapiens"}), returndf = TRUE)
temp[, -\hlfunctioncall{c}(1, 4)]
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
mynames <- \hlfunctioncall{c}(\hlstring{"Helianthus annuus"}, \hlstring{"Pinus contort"}, \hlstring{"Poa anua"}, \hlstring{"Abis magnifica"}, 
    \hlstring{"Rosa california"}, \hlstring{"Festuca arundinace"}, \hlstring{"Sorbus occidentalos"}, \hlstring{"Madia sateva"})
\hlfunctioncall{tnrs}(query = mynames)[, -\hlfunctioncall{c}(5:7)]
\end{alltt}
\begin{verbatim}
        submittedName        acceptedName    sourceId score
7   Helianthus annuus   Helianthus annuus iPlant_TNRS  1.00
4       Pinus contort      Pinus contorta iPlant_TNRS  0.98
5            Poa anua           Poa annua iPlant_TNRS  0.96
3      Abis magnifica     Abies magnifica iPlant_TNRS  0.96
8     Rosa california    Rosa californica iPlant_TNRS  0.99
2  Festuca arundinace Festuca arundinacea iPlant_TNRS  0.99
1 Sorbus occidentalos Sorbus occidentalis iPlant_TNRS  0.99
6        Madia sateva        Madia sativa iPlant_TNRS  0.97
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
specieslist <- \hlfunctioncall{c}(\hlstring{"Abies procera"}, \hlstring{"Pinus contorta"})
\hlfunctioncall{classification}(specieslist, db = \hlstring{"itis"})
\end{alltt}
\begin{verbatim}
[[1]]
        rankName       taxonName    tsn
1        Kingdom         Plantae 202422
2     Subkingdom  Viridaeplantae 846492
3   Infrakingdom    Streptophyta 846494
4       Division    Tracheophyta 846496
5    Subdivision Spermatophytina 846504
6  Infradivision    Gymnospermae 846506
7          Class       Pinopsida 500009
8          Order         Pinales 500028
9         Family        Pinaceae  18030
10         Genus           Abies  18031
11       Species   Abies procera 181835

[[2]]
        rankName       taxonName    tsn
1        Kingdom         Plantae 202422
2     Subkingdom  Viridaeplantae 846492
3   Infrakingdom    Streptophyta 846494
4       Division    Tracheophyta 846496
5    Subdivision Spermatophytina 846504
6  Infradivision    Gymnospermae 846506
7          Class       Pinopsida 500009
8          Order         Pinales 500028
9         Family        Pinaceae  18030
10         Genus           Pinus  18035
11       Species  Pinus contorta 183327
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{tax_name}(query = \hlstring{"Helianthus annuus"}, get = \hlstring{"family"}, db = \hlstring{"itis"})
\end{alltt}
\begin{verbatim}
      family
1 Asteraceae
\end{verbatim}
\begin{alltt}
\hlfunctioncall{tax_name}(query = \hlstring{"Helianthus annuus"}, get = \hlstring{"family"}, db = \hlstring{"ncbi"})
\end{alltt}
\begin{verbatim}
      family
1 Asteraceae
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{get_tsn}(searchterm = \hlstring{"Heliastes"}, searchtype = \hlstring{"sciname"})
\end{alltt}
\begin{verbatim}
           combinedname    tsn
1     Heliastes bicolor 615238
2   Heliastes chrysurus 615250
3     Heliastes cinctus 615573
4  Heliastes dimidiatus 615257
5  Heliastes hypsilepis 615273
6 Heliastes immaculatus 615639
7 Heliastes opercularis 615300
8      Heliastes ovalis 615301
 1 
NA 
attr(,"class")
[1] "tsn"
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
splist <- \hlfunctioncall{c}(\hlstring{"annona cherimola"}, \hlstring{"annona muricata"}, \hlstring{"quercus robur"}, \hlstring{"shorea robusta"}, 
    \hlstring{"pandanus patina"}, \hlstring{"oryza sativa"}, \hlstring{"durio zibethinus"})
\hlfunctioncall{get_tsn}(searchterm = splist, searchtype = \hlstring{"sciname"})
\end{alltt}
\begin{verbatim}
[1] "506198" "18098"  "19405"  "506787" "507376" "41976"  "506099"
attr(,"class")
[1] "tsn"
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{get_uid}(sciname = \hlfunctioncall{c}(\hlstring{"Chironomus riparius"}, \hlstring{"aaa vva"}))
\end{alltt}
\begin{verbatim}
[1] "315576" NA      
attr(,"class")
[1] "uid"
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
taxa <- \hlfunctioncall{c}(\hlstring{"Poa annua"}, \hlstring{"Abies procera"}, \hlstring{"Helianthus annuus"})
tree <- \hlfunctioncall{phylomatic_tree}(taxa = taxa)
tree$tip.label <- \hlfunctioncall{capwords}(tree$tip.label)
\hlfunctioncall{plot}(tree, cex = 1)
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{col_downstream}(name = \hlstring{"Apis"}, downto = \hlstring{"Species"})[[1]]
\end{alltt}
\begin{verbatim}
  childtaxa_id     childtaxa_name childtaxa_rank
1      6971712 Apis andreniformis        Species
2      6971713        Apis cerana        Species
3      6971714       Apis dorsata        Species
4      6971715        Apis florea        Species
5      6971716 Apis koschevnikovi        Species
6      6845885     Apis mellifera        Species
7      6971717   Apis nigrocincta        Species
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
ia <- \hlfunctioncall{iucn_summary}(\hlfunctioncall{c}(\hlstring{"Panthera uncia"}, \hlstring{"Lynx lynx"}))
\hlfunctioncall{iucn_status}(ia)
\end{alltt}
\begin{verbatim}
Panthera uncia      Lynx lynx 
          "EN"           "LC" 
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
out <- \hlfunctioncall{get_genes_avail}(taxon_name = \hlstring{"Umbra limi"}, seqrange = \hlstring{"1:2000"}, getrelated = FALSE)
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
out[\hlfunctioncall{grep}(\hlstring{"RAG1"}, out$genesavail, ignore.case = TRUE), ]
\end{alltt}
\begin{verbatim}
        spused length
413 Umbra limi    732
427 Umbra limi    959
434 Umbra limi   1631
                                                                            genesavail
413 isolate UlimA recombinase activating protein 1 (rag1) gene, exon 3 and partial cds
427           recombination-activating protein 1 (RAG1) gene, intron 2 and partial cds
434                        recombination-activating protein 1 (RAG1) gene, partial cds
    access_num       ids
413   JX190826 394772608
427   AY459526  45479841
434   AY380548  38858304
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
A <- \hlstring{"gammarus roeseli"}

B1 <- \hlstring{"gammarus roeseli"}
B2 <- \hlstring{"gammarus"}
B3 <- \hlstring{"gammaridae"}

A_clas <- \hlfunctioncall{classification}(A, db = \hlstring{'ncbi'})
B1_clas <- \hlfunctioncall{classification}(B1, db = \hlstring{'ncbi'})
B2_clas <- \hlfunctioncall{classification}(B2, db = \hlstring{'ncbi'})
B3_clas <- \hlfunctioncall{classification}(B3, db = \hlstring{'ncbi'})

B1[\hlfunctioncall{match}(A, B1)]
\end{alltt}
\begin{verbatim}
[1] "gammarus roeseli"
\end{verbatim}
\begin{alltt}
A_clas[[1]]$Rank[\hlfunctioncall{tolower}(A_clas[[1]]$ScientificName) %in% B2]
\end{alltt}
\begin{verbatim}
[1] "genus"
\end{verbatim}
\begin{alltt}
A_clas[[1]]$Rank[\hlfunctioncall{tolower}(A_clas[[1]]$ScientificName) %in% B3]
\end{alltt}
\begin{verbatim}
[1] "family"
\end{verbatim}
\end{kframe}
\end{knitrout}

If we find a direct match (here \emph{Gammarus roeseli}), we are lucky. But we can also match Gammaridae with \emph{Gammarus roeseli}, but on a lower taxonomic level. A more comprehensive and realistic example (matching a trait-table to a abundance table) is given in Appendix B.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{data}(dune, package = \hlstring{'vegan'})
df <- dune[ , 1:5]
\hlfunctioncall{colnames}(df) <- \hlfunctioncall{c}(\hlstring{"Bellis perennis"}, \hlstring{"Empetrum nigrum"}, \hlstring{"Juncus bufonius"}, 
                  \hlstring{"Juncus articulatus"}, \hlstring{"xxx"})
\hlfunctioncall{head}(df)
\end{alltt}
\begin{verbatim}
   Bellis perennis Empetrum nigrum Juncus bufonius Juncus articulatus xxx
2                3               0               0                  0   0
13               0               0               3                  0   0
4                2               0               0                  0   0
16               0               0               0                  3   0
6                0               0               0                  0   0
1                0               0               0                  0   0
\end{verbatim}
\begin{alltt}
agg <- \hlfunctioncall{tax_agg}(df, rank = \hlstring{'family'}, db = \hlstring{'ncbi'})
agg
\end{alltt}
\begin{verbatim}

@@

Level of Aggregation: FAMILY
No. taxa before aggregation: 5
No. taxa after aggregation: 4
No. taxa not found: 1
\end{verbatim}
\begin{alltt}
\hlfunctioncall{head}(agg$x)
\end{alltt}
\begin{verbatim}
   Asteraceae Ericaceae Juncaceae xxx
2           3         0         0   0
13          0         0         3   0
4           2         0         0   0
16          0         0         3   0
6           0         0         0   0
1           0         0         0   0
\end{verbatim}
\end{kframe}
\end{knitrout}

>>>>

\documentclass[10pt]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

@@

First, one must install and load taxize into the R session.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{install.packages}(\hlstr{"taxize"})
\hlkwd{library}(taxize)
\end{alltt}
\end{kframe}
\end{knitrout}

Advanced users can also download and install the latest development copy from GitHub \cite{github_taxize}.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
temp <- \hlkwd{gnr_resolve}(names = \hlkwd{c}(\hlstr{"Helianthos annus"}, \hlstr{"Homo saapiens"}), returndf = TRUE)
temp[, -\hlkwd{c}(1, 4)]
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
mynames <- \hlkwd{c}(\hlstr{"Helianthus annuus"}, \hlstr{"Pinus contort"}, \hlstr{"Poa anua"}, \hlstr{"Abis magnifica"}, 
    \hlstr{"Rosa california"}, \hlstr{"Festuca arundinace"}, \hlstr{"Sorbus occidentalos"}, \hlstr{"Madia sateva"})
\hlkwd{tnrs}(query = mynames)[, -\hlkwd{c}(5:7)]
\end{alltt}
\begin{verbatim}
        submittedName        acceptedName    sourceId score
7   Helianthus annuus   Helianthus annuus iPlant_TNRS  1.00
4       Pinus contort      Pinus contorta iPlant_TNRS  0.98
5            Poa anua           Poa annua iPlant_TNRS  0.96
3      Abis magnifica     Abies magnifica iPlant_TNRS  0.96
8     Rosa california    Rosa californica iPlant_TNRS  0.99
2  Festuca arundinace Festuca arundinacea iPlant_TNRS  0.99
1 Sorbus occidentalos Sorbus occidentalis iPlant_TNRS  0.99
6        Madia sateva        Madia sativa iPlant_TNRS  0.97
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
specieslist <- \hlkwd{c}(\hlstr{"Abies procera"}, \hlstr{"Pinus contorta"})
\hlkwd{classification}(specieslist, db = \hlstr{"itis"})
\end{alltt}
\begin{verbatim}
$`Abies procera`
        rankName       taxonName    tsn
1        Kingdom         Plantae 202422
2     Subkingdom  Viridaeplantae 846492
3   Infrakingdom    Streptophyta 846494
4       Division    Tracheophyta 846496
5    Subdivision Spermatophytina 846504
6  Infradivision    Gymnospermae 846506
7          Class       Pinopsida 500009
8          Order         Pinales 500028
9         Family        Pinaceae  18030
10         Genus           Abies  18031
11       Species   Abies procera 181835

$`Pinus contorta`
        rankName       taxonName    tsn
1        Kingdom         Plantae 202422
2     Subkingdom  Viridaeplantae 846492
3   Infrakingdom    Streptophyta 846494
4       Division    Tracheophyta 846496
5    Subdivision Spermatophytina 846504
6  Infradivision    Gymnospermae 846506
7          Class       Pinopsida 500009
8          Order         Pinales 500028
9         Family        Pinaceae  18030
10         Genus           Pinus  18035
11       Species  Pinus contorta 183327
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{tax_name}(query = \hlstr{"Helianthus annuus"}, get = \hlstr{"family"}, db = \hlstr{"itis"})
\end{alltt}
\begin{verbatim}
      family
1 Asteraceae
\end{verbatim}
\begin{alltt}
\hlkwd{tax_name}(query = \hlstr{"Helianthus annuus"}, get = \hlstr{"family"}, db = \hlstr{"ncbi"})
\end{alltt}
\begin{verbatim}
      family
1 Asteraceae
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{get_tsn}(searchterm = \hlstr{"Heliastes"}, searchtype = \hlstr{"sciname"})
\end{alltt}
\begin{verbatim}
           combinedname    tsn
1     Heliastes bicolor 615238
2   Heliastes chrysurus 615250
3     Heliastes cinctus 615573
4  Heliastes dimidiatus 615257
5  Heliastes hypsilepis 615273
6 Heliastes immaculatus 615639
7 Heliastes opercularis 615300
8      Heliastes ovalis 615301
 1 
NA 
attr(,"class")
[1] "tsn"
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
splist <- \hlkwd{c}(\hlstr{"annona cherimola"}, \hlstr{"annona muricata"}, \hlstr{"quercus robur"}, \hlstr{"shorea robusta"}, 
    \hlstr{"pandanus patina"}, \hlstr{"oryza sativa"}, \hlstr{"durio zibethinus"})
\hlkwd{get_tsn}(searchterm = splist, searchtype = \hlstr{"sciname"})
\end{alltt}
\begin{verbatim}
[1] "506198" "18098"  "19405"  "506787" "507376" "41976"  "506099"
attr(,"class")
[1] "tsn"
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{get_uid}(sciname = \hlkwd{c}(\hlstr{"Chironomus riparius"}, \hlstr{"aaa vva"}))
\end{alltt}
\begin{verbatim}
[1] "315576" NA      
attr(,"class")
[1] "uid"
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
taxa <- \hlkwd{c}(\hlstr{"Poa annua"}, \hlstr{"Abies procera"}, \hlstr{"Helianthus annuus"})
tree <- \hlkwd{phylomatic_tree}(taxa = taxa)
tree$tip.label <- \hlkwd{capwords}(tree$tip.label)
\hlkwd{plot}(tree, cex = 1)
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{col_downstream}(name = \hlstr{"Apis"}, downto = \hlstr{"Species"})[[1]]
\end{alltt}
\begin{verbatim}
  childtaxa_id     childtaxa_name childtaxa_rank
1      6971712 Apis andreniformis        Species
2      6971713        Apis cerana        Species
3      6971714       Apis dorsata        Species
4      6971715        Apis florea        Species
5      6971716 Apis koschevnikovi        Species
6      6845885     Apis mellifera        Species
7      6971717   Apis nigrocincta        Species
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
ia <- \hlkwd{iucn_summary}(\hlkwd{c}(\hlstr{"Panthera uncia"}, \hlstr{"Lynx lynx"}))
\hlkwd{iucn_status}(ia)
\end{alltt}
\begin{verbatim}
Panthera uncia      Lynx lynx 
          "EN"           "LC" 
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
out <- \hlkwd{get_genes_avail}(taxon_name = \hlstr{"Umbra limi"}, seqrange = \hlstr{"1:2000"}, getrelated = FALSE)
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
out[\hlkwd{grep}(\hlstr{"RAG1"}, out$genesavail, ignore.case = TRUE), ]
\end{alltt}
\begin{verbatim}
        spused length
413 Umbra limi    732
427 Umbra limi    959
434 Umbra limi   1631
                                                                            genesavail
413 isolate UlimA recombinase activating protein 1 (rag1) gene, exon 3 and partial cds
427           recombination-activating protein 1 (RAG1) gene, intron 2 and partial cds
434                        recombination-activating protein 1 (RAG1) gene, partial cds
    access_num       ids
413   JX190826 394772608
427   AY459526  45479841
434   AY380548  38858304
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
A <- \hlstr{"gammarus roeseli"}

B1 <- \hlstr{"gammarus roeseli"}
B2 <- \hlstr{"gammarus"}
B3 <- \hlstr{"gammaridae"}

A_clas <- \hlkwd{classification}(A, db = \hlstr{'ncbi'})
B1_clas <- \hlkwd{classification}(B1, db = \hlstr{'ncbi'})
B2_clas <- \hlkwd{classification}(B2, db = \hlstr{'ncbi'})
B3_clas <- \hlkwd{classification}(B3, db = \hlstr{'ncbi'})

B1[\hlkwd{match}(A, B1)]
\end{alltt}
\begin{verbatim}
[1] "gammarus roeseli"
\end{verbatim}
\begin{alltt}
A_clas[[1]]$Rank[\hlkwd{tolower}(A_clas[[1]]$ScientificName) %in% B2]
\end{alltt}
\begin{verbatim}
[1] "genus"
\end{verbatim}
\begin{alltt}
A_clas[[1]]$Rank[\hlkwd{tolower}(A_clas[[1]]$ScientificName) %in% B3]
\end{alltt}
\begin{verbatim}
[1] "family"
\end{verbatim}
\end{kframe}
\end{knitrout}

If we find a direct match (here \emph{Gammarus roeseli}), we are lucky. But we can also match Gammaridae with \emph{Gammarus roeseli}, but on a lower taxonomic level. A more comprehensive and realistic example (matching a trait table with an abundance table) is given in Appendix B.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{data}(dune, package = \hlstr{'vegan'})
df <- dune[ , 1:5]
\hlkwd{colnames}(df) <- \hlkwd{c}(\hlstr{"Bellis perennis"}, \hlstr{"Empetrum nigrum"}, \hlstr{"Juncus bufonius"}, 
                  \hlstr{"Juncus articulatus"}, \hlstr{"xxx"})
\hlkwd{head}(df)
\end{alltt}
\begin{verbatim}
   Bellis perennis Empetrum nigrum Juncus bufonius Juncus articulatus xxx
2                3               0               0                  0   0
13               0               0               3                  0   0
4                2               0               0                  0   0
16               0               0               0                  3   0
6                0               0               0                  0   0
1                0               0               0                  0   0
\end{verbatim}
\begin{alltt}
agg <- \hlkwd{tax_agg}(df, rank = \hlstr{'family'}, db = \hlstr{'ncbi'})
agg
\end{alltt}
\begin{verbatim}

@@

Level of Aggregation: FAMILY
No. taxa before aggregation: 5
No. taxa after aggregation: 4
No. taxa not found: 1
\end{verbatim}
\begin{alltt}
\hlkwd{head}(agg$x)
\end{alltt}
\begin{verbatim}
   Asteraceae Ericaceae Juncaceae xxx
2           3         0         0   0
13          0         0         3   0
4           2         0         0   0
16          0         0         3   0
6           0         0         0   0
1           0         0         0   0
\end{verbatim}
\end{kframe}
\end{knitrout}
